\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-11-29 smurf_paper.tex
add section about lags to pre-processing

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
thermal time constant, as well as a negative shift of about $-$2.5\,ms
caused by a half-sample offset due to the synchronization between the
pointing and bolometer data, yields a net lag of approximately
4.5\,ms. There are also line features in the spectra that are thought
to be produced primarily at frequencies far above the final 200\,Hz
sample rate, and are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

@@

In principle there are two other effects that could be included in
this model. First, signals detected by the bolometers (inside the
brackets of Eq.~\ref{eq:model}) should be convolved with the thermal
response function, $\exp(-t/\tau)$, where $\tau=1$\,ms. Second, the
final signals should be convolved with the response of the
anti-aliasing filter. In practice, the combined effect of these two
filters is to introduce a lag of $\sim$7\,ms, and only minimal
attentuation of a point source (at most 3\% for the smaller
450\,\micron\ PSF at the maximum scan speed of 600\,arcsec). For all
of the maps described in this paper, their impact is therefore
negligible. The current version of SMURF (part of the Starlink
\textsc{kapuahi} release) does not attempt to de-convolve these
effects, although the next version will likely include an optional
pre-processing step to perform this task.

@@

For maps of faint point sources, a single (non-iterative) high-pass
filter at the start of the reduction produces maps that are nearly
white-noise limited and linear (i.e., the response does not depend on
\snr). Residual large-scale noise can be removed with a whitening
filter (also established from jackknife estimates of the noise) based
on the Fourier Transform of the maps, and sources detected using a
matched-filter (smoothing with the effective filtered point spread
function).

>>>>

In addition to uncorrelated white noise, and low-frequency (and often
correlated from bolometer to bolometer) $1/f$ drifts, bolometer power
spectra exhibit a roll-off at frequencies approaching Nyquist, which
is due to the anti-aliasing filter that forms part of the 200\,Hz
re-sampling stage. This anti-aliasing filter also introduces an
effective lag of $\sim$6\,ms. Adding this lag to the $\sim$1\,ms
thermal time constant, as well as a negative shift of about $-$2.5\,ms
caused by a half-sample offset due to the synchronization between the
pointing and bolometer data, yields a net lag of approximately
4.5\,ms. There are also line features in the spectra that are thought
to be produced primarily at frequencies far above the final 200\,Hz
sample rate, which are aliased to lower frequencies during the
multiplexed readout stage before they can be removed by the
anti-aliasing filter.

@@

\subsubsection{Impact of lags}
\label{sec:lags}

In principle, the effects of the bolometer thermal response,
anti-aliasing filter, and data acquisition lags should be accounted
for; together they produce a net lag in the bolometer signals of about
4.5\,ms (Section~\ref{sec:bolos}). This can be modelled by convolving
the ``pure'' bolometer signals with a single system response
function. Similarly, this effect could be removed using de-convolution
as a pre-processing step. The impact of this lag is primarily a
reduction in point-source sensitivity since scans at different
position angles will detect the source at slightly different positions
(and therefore reduce the peak signal once averaged together). This
attenuation is most significant at 450\,\micron\ given the smaller
beam size, and at the greatest scan speeds. For reference, at a speed
of 150\,arcsec\,s$^{-1}$ (typical for CV daisies of calibrators and
deep fields) the attenuation is a negligible $<$1\% at 850\,\micron,
and $\sim$1\% at 450\,\micron. However, at the maximum speed of
600\,arcsec\,s$^{-1}$ (for the largest, and typically shallowest
maps), this attenuation grows to about 5\% and 15\% at 850 and
450\,\micron, respectively. It is unlikely that this response has had
a significant impact on observations taken to date, since such scans
have been used primarily for shallow cosmology fields for which only
the 850\,\micron\ data are expected to be useful, or Galactic fields
of extended structures (which are less sensitive to this effect). In
the present Starlink \textsc{kapuahi} release there is no option to
de-convolve the system response. However, it is presently being added
and should be present in future releases.

@@

@@

For maps of faint point-sources, a single (non-iterative) high-pass
filter at the start of the reduction produces maps that are nearly
white-noise limited and linear (i.e., the response does not depend on
\snr). Residual large-scale noise can be removed with a whitening
filter (also established from jackknife estimates of the noise) based
on the Fourier Transform of the maps, and sources detected using a
matched-filter (smoothing with the effective filtered point spread
function).
